$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-05 paper/paper.tex
Use different counters for definitions, lemmas, and theorems
Name our theorems

% Do not use italics in definitions and theorems
\theoremstyle{definition}
\newtheorem{nidefinition}[theorem]{Definition}
\newtheorem{nitheorem}[theorem]{Theorem}
\newtheorem{nilemma}[theorem]{Lemma}

@@

\begin{nitheorem}
  \label{thm:framing}
  Framing captures the idea that the well-typedness of a process is independent of its leftover resources.
  Let $P$ be a well-typed process $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
  Let there be a $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
  Take arbitrary contexts $\Gamma_r$ and $\Xi_r$ such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
\end{nitheorem}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

\begin{nitheorem}
  \label{thm:weakening}
  Weakening captures the idea that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.

@@

\begin{nitheorem}
  \label{thm:strengthening}
  Strengthening captures the idea that removing an unused variable preserves the well-typedness of a process.

@@

\begin{nitheorem}
  \label{thm:swapping}
  Swapping two variables preserves the well-typedness of a process.
  Let $\func{swap}_i$ be a function that swaps variables $i$ and $\suc i$.
  For simplicity, we overload $\func{swap}$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
  \todo{}
  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
\end{nitheorem}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

\begin{nitheorem}
  \label{thm:subject-congruence}
  Applying structural congruence rules to a well-typed process preserves its well-typedness.
  That is, if $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
\end{nitheorem}

@@

\begin{nitheorem}
  \label{thm:subst-generalization}
  Let process $P$ be well-typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  The substituted variable $i$ has a usage $m$ in $\Gamma_i$, and a usage $n$ in $\Psi_i$.
  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
  In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
  \begin{multicols}{2}
  \begin{itemize}
    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
  \end{itemize}
  \end{multicols}
  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
  Let $\Delta$ have a usage annotation $\lz$ at position $i$.
  Then substituting $i$ with $j$ in $P$ will be well-typed such that $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
  Refer to \autoref{fig:subst} for a diagramatic representation.
\end{nitheorem}

@@

\begin{nitheorem}
  \label{thm:substitution}
  Let $P$ be a process typed under $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
  Let $\ni_j m$ be a \textsc{VarRef} $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well-typed under $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
\end{nitheorem}
\begin{proof}
  Use framing to derive $\contains{\gamma}{\Gamma}{j}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
  Supply these to the generalised \autoref{thm:subst-generalization}.
\end{proof}

@@

\begin{nitheorem}
  Let $P \reduce{c} Q$.
  If $c$ is $\constr{internal}$, then $\Gamma' \equiv \Gamma$.
  If however $c$ is $\constr{external}\; i$, then $\containsusage{\Gamma'}{i}{\lio}{\Gamma}$.
  Given $\types{\gamma}{\Gamma'}{P}{\Xi}$ then after reduction $\types{\gamma}{\Gamma}{Q}{\Xi}$.
\end{nitheorem}

>>>>

% Do not use italics in definitions and theorems
\theoremstyle{definition}
\newtheorem{nidefinition}{Definition}
\newtheorem{nitheorem}{Theorem}
\newtheorem{nilemma}{Lemma}

@@

\begin{nitheorem}[Framing]
  \label{thm:framing}
  The well-typedness of a process is independent of its leftover resources.
  Let $P$ be a well-typed process $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
  Let there be a $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
  Take arbitrary contexts $\Gamma_r$ and $\Xi_r$ such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
\end{nitheorem}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

\begin{nitheorem}[Weakening]
  \label{thm:weakening}
  Inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.

@@

\begin{nitheorem}[Strengthening]
  \label{thm:strengthening}
  Removing an unused variable preserves the well-typedness of a process.

@@

\begin{nitheorem}[Swapping]
  \label{thm:swapping}
  Swapping two variables preserves the well-typedness of a process.
  Let $\func{swap}_i$ be a function that swaps variables $i$ and $\suc i$.
  For simplicity, we overload $\func{swap}$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
  \todo{}
  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
\end{nitheorem}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

@@

\begin{nitheorem}[Subject congruence]
  \label{thm:subject-congruence}
  Applying structural congruence rules to a well-typed process preserves its well-typedness.
  That is, if $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
\end{nitheorem}

@@

\begin{nitheorem}[Generalised substitution]
  \label{thm:subst-generalization}
  Let process $P$ be well-typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  The substituted variable $i$ has a usage $m$ in $\Gamma_i$, and a usage $n$ in $\Psi_i$.
  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
  In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
  \begin{multicols}{2}
  \begin{itemize}
    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
  \end{itemize}
  \end{multicols}
  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
  Let $\Delta$ have a usage annotation $\lz$ at position $i$.
  Then substituting $i$ with $j$ in $P$ will be well-typed such that $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
  Refer to \autoref{fig:subst} for a diagramatic representation.
\end{nitheorem}

@@

\begin{nitheorem}[Substitution]
  \label{thm:substitution}
  Let $P$ be a process typed under $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
  Let $\ni_j m$ be a \textsc{VarRef} $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well-typed under $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
\end{nitheorem}
\begin{proof}
  Use framing to derive $\contains{\gamma}{\Gamma}{j}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
  Supply these to the generalised \autoref{thm:subst-generalization}.
\end{proof}

@@

\begin{nitheorem}[Subject reduction]
  Let $P \reduce{c} Q$.
  If $c$ is $\constr{internal}$, then $\Gamma' \equiv \Gamma$.
  If however $c$ is $\constr{external}\; i$, then $\containsusage{\Gamma'}{i}{\lio}{\Gamma}$.
  Given $\types{\gamma}{\Gamma'}{P}{\Xi}$ then after reduction $\types{\gamma}{\Gamma}{Q}{\Xi}$.
\end{nitheorem}
